---
title: How To Show Mean Value in Boxplots with ggplot2"
author: "Craig W. Slinkman"
date: "10/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question

Titanium wants an answer to:

> Which would best display the following data if you wanted to display the numbers which are outliers as well as the mean?	
  
## Answer

In my opinion the best approach use to draw a box plot *called a box-and-whisker plot by proposed John tukey Tukey (https://asq.org/quality-resources/quality-glossary/b). We will demonstrate the used of boxplot by using $R$'s $ggplot2$ package by using $ggplot2$ package to drawn a boxplot.

The data set that we will use is called $gapminder$. The $gapminder$ data set contains data about life expectancy, GDP per capita, and population for the country, spanning many decades.  This example is based on
https://datavizpyr.com/how-to-add-mean-mark-to-boxplot-in-ggplot2/.

I assume that uou hve the required packages installed.  If not you can install these packages by using the
the statements below to install the packages in R.

```
install.packages( "tidyverse", "gapminder")
```



 
We start by telling $R$ that we are going to use $R$'s $tidyverse$ package.  This automate loads the graphics package $ggplot2$.

```{r}
library( tidyverse )                    # I live in the tidyverse.
library( gapminder )                    # Provides the gapminder data. 
library( cowplot )                      # For publishable graphics.
```

We verify that that the gapminder data is available for use br printing the tibble:

```{r}
gapminder
```

We start by drawing a boxplot.

```{r}
p1 <-                                       # Define plot
    gapminder %>%                     
        ggplot(aes( continent,
                   lifeExp, 
                    fill=continent)) +
            geom_boxplot() +
            theme_cowplot() +
         theme(legend.position = "none")
p1                                          # Print plot
```

## Adding the mean to the boxplots

We add the means to the boxplots by using the following statements:

```{r}
p2 <- p1 + stat_summary(fun.y="mean")
p2
```
Note that the means aooear as black circles.  

## Analysis

We see that each distribution is approximately symetric.  There is a very pronounced difference distributions in the value of the central tendency and dispersion across continents.

## We save the plots for bublication.

```{r}
save_plot( "boxplot_with_mean.png"
           p2 )
```

