---
title: "Geometric distribution question"
author: "Craig Slinkman"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


My interpretation of your question. (Remark use full sentences to write your question.  It will save everybody time)

> Suppose we have a geometriuc distribution with a variabnce of 2.0.  Compute the probability of $Pr( 5 \le X \le 8 ).probability 

We will answer this question in three parts:

* Computing the probability of success from the variance.
* Computing the answer using the probability mass function.
* Computing the probability using the cumulative distribution function.

### Computing the probability of success from the variance

We are given the variance.  The equation for computing the variance of a geometric distribution with parameter $p$ is 

$$Var(X) = \frac{1-p}{p^2}$$

Thus we need to solve this equation when $Var(X)=2$.  Hence we have:

$$\frac{1-p}{p^2} = 2$$

We multiply both sides by $p^2$ to obtain'
'
$$1-p = 2 p^2$$
Subtract $-2p^2$ from both sides:

$$-2 p^2 -p + 1 = 0$$

We can factor the above equation to get

$$-(2p - 1)(p+1)=0$$
Therefore we obtain two equations:

$$$2p-1=0\$$
and 
$$p+1 = 0$$
Which gives the following solution set

Now since $p$ is the probability of success and must be between zero and 1.  The answer to that $p = 1/2$. 

We check this by substituting into the equation for the variance of the geometric distribution:

$$Var(X|p=0.5) =  \frac{1-p}{p^2} = \frac{1-0.5}{0.5^2} = \frac{0.5}{0.25} = 2$$
So the solution to the quadratic question is 

Now we must compute the individual probabilities using the geometric distribution's probability mass function.

The equation for the probability mass function is:

$$Pr(X=k|p) = p(1-p)^{k-1}$$
We have $p=0.5$, Substituting this value in the above equation we have:

Pr(X=k|p=0.5) = 0.5 \cdot (1-0.5)^{k-1} = p^k$$

```{r}
p5 <- 0.5^5
p6 <- 0.5^6
p7 <- 0.5^7
p8 <- 0.5^8

p58 = p5 + p6 + p7 +p8

```


Now we can compute the individual probabilities:

$$Pr(X=5|p=0.5)`= `r p5`$$
$$Pr(x=6|p=0.5) = `r p6`$$
$$Pr(x=7|p=0.5)  = `r p7`$$
$$Pr(x= 8|p-0.5)   = `r p8`$$

Thus we have:
$$Pr(5 \le X \le 8) = `r p5` + `r p6` + 'r p7` + `r p8` = `r p58`$$

### Computing the answer using the cumulative distribution function


 
```

